---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readr)
```

inclusion/exclusion criteria
1. at least 1 encounter in 2012 - 2014
I specifically defined 1 encounter as 1 record in the encounter diagnosis table or the billing table
```{r}
Billing <- read_csv("~/Documents/Grad School 2018 Fall/ASD/Billing.csv", 
    col_types = cols_only(DiagnosisCodeType_calc = col_guess(), 
        DiagnosisCode_calc = col_guess(), 
        DiagnosisText_calc = col_guess(), 
        Network_ID = col_guess(), Patient_ID = col_character(), 
        ServiceCode = col_guess(), ServiceDate = col_guess(), 
        Site_ID = col_character()))


Billing['Year'] <- substring(Billing$ServiceDate, 1, 4)

billing <- Billing %>% filter(Year == "2012" | Year == "2013" | Year == "2014")

#unique(c(unique(billing$Patient_ID), unique(ed$Patient_ID)))
EncounterDiagnosis <- read_csv("~/Documents/Grad School 2018 Fall/ASD/EncounterDiagnosis.csv", 
    col_types = cols_only(DateCreated = col_guess(), 
        DiagnosisCode_calc = col_guess(), 
        DiagnosisText_calc = col_guess(), 
        Network_ID = col_guess(), Patient_ID = col_character(), 
        Site_ID = col_character()))


EncounterDiagnosis['Year'] <- substring(EncounterDiagnosis$DateCreated, 1, 4)
ed <- EncounterDiagnosis %>% filter(Year == "2012" | Year == "2013" | Year == "2014")

one <- tbl_df(unique(c(unique(billing$Patient_ID), unique(ed$Patient_ID))))
#wanted to get a list of patients that exist in billing table and encounter diagnosis table

Patient <- read_csv("~/Documents/Grad School 2018 Fall/ASD/Patient.csv", 
    col_types = cols_only(BirthYear = col_number(), 
        Patient_ID = col_character(), Sex = col_guess()))

one['Age'] <- NA
one$Age <- 2016 - Patient$BirthYear[match(one$value, Patient$Patient_ID)]

two <- one %>% filter(Age<= 25)

b <- Billing[Billing$Patient_ID %in% one, ] 
ed <- EncounterDiagnosis[EncounterDiagnosis$Patient_ID %in% one, ]

#one is the cohort of all the people who have at least 1 encounter in 2012 - 2014
#b is billing records for all years for all patients who were active in 2012 - 2014
#ed is encounter diagnosis records for all years for all patients who were active in 2012 - 2014
```

2 physician visits with ASD ICD code or ASD in problem list
billing code = 299

These must have occured from the years of 2012 - 2014? And then after their first ASD visit what happened? 

Does it have to be ASD visit or first ASD visit? 

For all the people in cohort, when was hte first time they had ASD? it's likely outside of our study range
```{r}
HealthCondition <- read_csv("~/Documents/Grad School 2018 Fall/ASD/HealthCondition.csv", 
    col_types = cols_only(ActiveInactiveFlag = col_guess(), 
        DateCreated = col_guess(), DateOfOnset = col_guess(), 
        DiagnosisCodeType_calc = col_guess(), 
        DiagnosisCode_calc = col_guess(), 
        DiagnosisText_calc = col_guess(), 
        Patient_ID = col_character(), SignificantNegativeFlag = col_guess(), 
        Site_ID = col_character()))

asd.hc <- HealthCondition %>% filter(substring(DiagnosisCode_calc, 1, 3) == '299') %>% mutate(Year = substring(DateCreated, 1, 4))


asd.b <- Billing %>% filter(substring(DiagnosisCode_calc, 1, 3) == '299')

asd.ed <- EncounterDiagnosis %>% filter(substring(DiagnosisCode_calc, 1, 3) == "299")

asd.visits <- full_join(asd.b %>% group_by(Patient_ID) %>% summarize(billing_n = n()), asd.ed %>% group_by(Patient_ID) %>% summarize(ed_n = n()))

asd.visits[is.na(asd.visits)] <- 0
asd.visits['sum_n'] <- rowSums(asd.visits[2:3], )

three <- unique(c(asd.visits$Patient_ID[asd.visits$sum_n > 1], asd.hc$Patient_ID)) # at least two asd visits, 

# two.2 <- unique(c(asd.visits$Patient_ID[asd.visits$sum_n > 0], asd.hc$Patient_ID))
# two.2 <- tbl_df(two.2[two.2 %in% one])

three <- tbl_df(three[three %in% two$value])

#two is the cohort of people who have at least 2 asd encounters or 1 asd record in health condition in 2012-2014, and who have encounters in 2012 - 2014
#previously two is 3110 rows
#now two is 1281
```

subset of two, when is the first time that these people had ASD related visits? 
take everyone who has had billing code asd (299) from any of the 3 tables, 
```{r}
firstasd.b <- Billing %>% filter(substring(DiagnosisCode_calc, 1, 3) == '299') %>% group_by(Patient_ID) %>% slice(which.min(ymd(ServiceDate))) %>% select(Patient_ID, ServiceDate)

firstasd.e <- EncounterDiagnosis %>% filter(substring(DiagnosisCode_calc, 1, 3) == '299') %>% group_by(Patient_ID) %>% slice(which.min(ymd(DateCreated))) %>% select(Patient_ID, DateCreated)

firstasd.hc <- HealthCondition %>% filter(substring(DiagnosisCode_calc, 1, 3) == '299') %>% group_by(Patient_ID) %>% slice(which.min(ymd(DateCreated))) %>% select(Patient_ID, DateCreated)

firstasd <- full_join(full_join(firstasd.b, firstasd.e, by = "Patient_ID"), firstasd.hc, by = "Patient_ID")

firstasd['earliestdate'] <- apply(firstasd[,2:4], 1, min, na.rm = T)

one['firstasd'] <- NA
one$firstasd <- firstasd$earliestdate[match(one$value, firstasd$Patient_ID)]

first ASD not important as long as they have ASD at some point
# three <- one %>% filter(substring(firstasd, 1, 4) == "2012" |substring(firstasd, 1, 4) == "2013" | substring(firstasd, 1, 4) == "2012")
# 
# three['Patient_ID'] <- three$value

# two.2['firstasd'] <- firstasd$earliestdate[match(two.2$value, firstasd$Patient_ID)]
# 
# two.2 <- two.2 %>% mutate(Year = substring(firstasd, 1, 4)) %>% filter(Year >= 2012)

#firstasd contains patient IDs and the first ever recorded ASD billing code dates for all patients who has ever had ASD in dataset

#note that firstasd DOES NOT include people who had first ASD AFTER 2014
```

matched cohort - need to have no neurological disorders
```{r}
#find list of patients who have never had these codes in either billing table, health condition table, or encounter daignosis table


neuro.hc <- HealthCondition %>% filter(substring(DiagnosisCode_calc, 1, 3) == '345' | DiagnosisCode_calc == "780.39" | substring(DiagnosisCode_calc, 1, 3) == "758" | DiagnosisCode_calc == "277.00" | substring(DiagnosisCode_calc, 1, 5) == "333.4" | substring(DiagnosisCode_calc, 1, 5) == "282.6" | substring(DiagnosisCode_calc, 1, 5) == "759.8" | substring(DiagnosisCode_calc, 1, 5) == "299.0" | DiagnosisCode_calc == "299.80" |   substring(DiagnosisCode_calc, 1, 5) == "299.9" | DiagnosisCode_calc == "314.00" | substring(DiagnosisCode_calc, 1, 5) == "343.9" | substring(DiagnosisCode_calc, 1, 3) == '317' | substring(DiagnosisCode_calc, 1, 3) == '318' | substring(DiagnosisCode_calc, 1, 3) == '319' |    
DiagnosisCode_calc == "760.71" | DiagnosisCode_calc == "307.22" | DiagnosisCode_calc == "307.23" |  substring(DiagnosisCode_calc, 1, 3) == "315")

neuro.b <- Billing %>% filter(substring(DiagnosisCode_calc, 1, 3) == '345' | DiagnosisCode_calc == "780.39" | substring(DiagnosisCode_calc, 1, 3) == "758" | DiagnosisCode_calc == "277.00" | substring(DiagnosisCode_calc, 1, 5) == "333.4" | substring(DiagnosisCode_calc, 1, 5) == "282.6" | substring(DiagnosisCode_calc, 1, 5) == "759.8" | substring(DiagnosisCode_calc, 1, 5) == "299.0" | DiagnosisCode_calc == "299.80" |   substring(DiagnosisCode_calc, 1, 5) == "299.9" | DiagnosisCode_calc == "314.00" | substring(DiagnosisCode_calc, 1, 5) == "343.9" | substring(DiagnosisCode_calc, 1, 3) == '317' | substring(DiagnosisCode_calc, 1, 3) == '318' | substring(DiagnosisCode_calc, 1, 3) == '319' |    
DiagnosisCode_calc == "760.71" | DiagnosisCode_calc == "307.22" | DiagnosisCode_calc == "307.23" |  substring(DiagnosisCode_calc, 1, 3) == "315")

neuro.ed <- EncounterDiagnosis %>% filter(substring(DiagnosisCode_calc, 1, 3) == '345' | DiagnosisCode_calc == "780.39" | substring(DiagnosisCode_calc, 1, 3) == "758" | DiagnosisCode_calc == "277.00" | substring(DiagnosisCode_calc, 1, 5) == "333.4" | substring(DiagnosisCode_calc, 1, 5) == "282.6" | substring(DiagnosisCode_calc, 1, 5) == "759.8" | substring(DiagnosisCode_calc, 1, 5) == "299.0" | DiagnosisCode_calc == "299.80" |   substring(DiagnosisCode_calc, 1, 5) == "299.9" | DiagnosisCode_calc == "314.00" | substring(DiagnosisCode_calc, 1, 5) == "343.9" | substring(DiagnosisCode_calc, 1, 3) == '317' | substring(DiagnosisCode_calc, 1, 3) == '318' | substring(DiagnosisCode_calc, 1, 3) == '319' |    
DiagnosisCode_calc == "760.71" | DiagnosisCode_calc == "307.22" | DiagnosisCode_calc == "307.23" |  substring(DiagnosisCode_calc, 1, 3) == "315")


neuro.vec <- unique(c(unique(neuro.hc$Patient_ID), unique(neuro.b$Patient_ID), unique(neuro.ed$Patient_ID)))

three <- tbl_df(three[three %in% two$value])
```

want to ensure that all patients have at least records in medication table
```{r}
Medication <- read_csv("~/Documents/Grad School 2018 Fall/ASD/Medication.csv", 
    col_types = cols_only(Code_calc = col_guess(), 
        DateCreated = col_guess(), Patient_ID = col_character(), 
        StartDate = col_guess(), StopDate = col_guess()))

AllergyIntolerance <- read_csv("~/Documents/Grad School 2018 Fall/ASD/AllergyIntolerance.csv", 
    col_types = cols_only(AllergyStatus = col_character(), 
        Category = col_guess(), CodeType_calc = col_guess(), 
        Code_calc = col_guess(), DateCreated = col_guess(), 
        Name_calc = col_guess(), Patient_ID = col_character(), 
        ReactionType = col_character(), Severity = col_character(), 
        StartDate = col_guess()))

Referral <- read_csv("~/Documents/Grad School 2018 Fall/ASD/Referral.csv", 
    col_types = cols_only(DateCreated = col_guess(), 
        Name_calc = col_guess(), Patient_ID = col_character()))

PatientProvider <- read_csv("~/Documents/Grad School 2018 Fall/ASD/PatientProvider.csv", 
    col_types = cols_only(Patient_ID = col_character(), 
        Provider_ID = col_character()))

Provider <- read_csv("~/Documents/Grad School 2018 Fall/ASD/Provider.csv", 
    col_types = cols(Provider_ID = col_character()))

Network <- read_csv("~/Documents/Grad School 2018 Fall/ASD/Network.csv")

Lab <- read_csv("~/Documents/Grad School 2018 Fall/ASD/Lab.csv", 
    col_types = cols_only(Name_calc = col_character(), 
        Patient_ID = col_character(), PerformedDate = col_character(), 
        TestResult_calc = col_guess(), UnitOfMeasure_calc = col_guess()))

HealthCondition <- read_csv("~/Documents/Grad School 2018 Fall/ASD/HealthCondition.csv", 
    col_types = cols_only(ActiveInactiveFlag = col_guess(), 
        DateOfOnset = col_guess(), DiagnosisCode_calc = col_guess(), 
        DiagnosisText_calc = col_guess(), 
        Patient_ID = col_character()))


Exam <- read_csv("~/Documents/Grad School 2018 Fall/ASD/Exam.csv", 
    col_types = cols_only(DateCreated = col_guess(), 
        Exam1 = col_guess(), Exam2 = col_character(), 
        Patient_ID = col_character(), Result1_calc = col_guess(), 
        Result2_calc = col_double(), UnitOfMeasure_calc = col_guess()))



med <- Medication %>% filter(!is.na(DateCreated)) %>% filter(substring(DateCreated, 1, 4) >= 2012)

DiseaseCase <- read_csv("~/Documents/Grad School 2018 Fall/ASD/DiseaseCase.csv", 
    col_types = cols_only(DateOfOnset = col_guess(), 
        Disease = col_guess(), Patient_ID = col_character()))

cohort <- three[three$Patient_ID %in% med$Patient_ID, ]

billing <- Billing[Billing$Patient_ID %in% cohort$Patient_ID, ]
encounterdiagnosis <- EncounterDiagnosis[EncounterDiagnosis$Patient_ID %in% cohort$Patient_ID, ]
medication <- Medication[Medication$Patient_ID %in% cohort$Patient_ID, ]
allergyintolerance <- AllergyIntolerance[AllergyIntolerance$Patient_ID %in% cohort$Patient_ID, ]
referral <- Referral[Referral$Patient_ID %in% cohort$Patient_ID, ]
lab <- Lab[Lab$Patient_ID %in% cohort$Patient_ID, ]
healthcondition <- HealthCondition[HealthCondition$Patient_ID %in% cohort$Patient_ID, ]

exam <- Exam[Exam$Patient_ID %in% cohort$Patient_ID, ]

diseasecase <- DiseaseCase[DiseaseCase$Patient_ID %in%  cohort$Patient_ID, ]
```

```{r}
two.2['encounterdate'] <- NA

two.2['Patient_ID'] <- two.2$value
#names(two.2) <- c("Patient_ID", "EncounterDate")
ed.two.2 <- ed[ed$Patient_ID %in% two.2$Patient_ID, ]
b.two.2 <- b[b$Patient_ID %in% two.2$Patient_ID,]

#encounterdiagnosis 
three.2 <- merge(ed.two.2 %>% group_by(Patient_ID) %>% slice(which.max(ymd(DateCreated))) %>% mutate(eddate = DateCreated) %>% select(Patient_ID, eddate), b.two.2 %>% group_by(Patient_ID) %>% slice(which.max(ymd(ServiceDate))) %>% mutate(billingdate = ServiceDate) %>%  select(Patient_ID, billingdate))

three.2['date'] <- pmax(three.2$billingdate, three.2$eddate)



three.2['age'] <- as.numeric(substring(three.2$date, 1, 4)) - as.numeric(Patient$BirthYear[match(three.2$Patient_ID, Patient$Patient_ID)])

#cohort 
cohort.2 <- three.2 %>% filter(age <= 18)
table(substring(cohort.2$date, 1, 4))

med.2 <- Medication[Medication$Patient_ID %in% cohort.2$Patient_ID, ]
med.2['firstdate'] <- firstasd$earliestdate[match(med.2$Patient_ID, firstasd$Patient_ID)]

med.2 <- med.2 %>% filter(DateCreated >= firstdate)
length(unique(med.2$Patient_ID))

#length of other cohort is 391
```
